---
title: "Assembling supplementary figures and tables" 
author: Laura Symul
format: 
   html:
     page-layout: full
     code-fold: true
     toc: true
     toc-location: left
     toc-depth: 5
     embed-resources: true
execute:
  cache: refresh  # true | refresh
knitr:
  opts_chunk:
    out.width: "100%"
editor: source
---


```{r}
#| echo: false
#| warning: false
#| cache: false

rm(list = ls())

library(tidyverse)
library(magrittr)
library(patchwork)
library(gt)
library(ggtext)


```


```{r}
#| echo: false
#| warning: false
#| cache: false

tmp <- fs::dir_map("../R/", source)

```

```{r}
#| echo: false
#| warning: false
#| cache: false

theme_set(theme_publication())

```


```{r}
#| warning: false

# We load the MAE object
mae <- load_latest_mae(dir = str_c(data_dir(), "03_augmented_mae/")) 
clin <- MultiAssayExperiment::colData(mae) |> as.data.frame() |> as_tibble()

```




## [deprecated] Supplementary Figure 1 (before revisisons)


```{r}

load(str_c(fig_out_dir(), "supplementary_ggplots/S1A.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S1B.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S1C.RData"), verbose = TRUE)

```

```{r}
#| fig-width: 18
#| fig-height: 8

g_S1B[[1]] <- g_S1B[[1]] + labs(tag = "B")
g_S1C[[1]] <- g_S1C[[1]] + labs(tag = "C")

fig_S1 <- 
  free(g_S1A + labs(tag = "A")) + # + theme(legend.position = "bottom")
  g_S1B +
  g_S1C + 
  plot_layout(
    widths = c(1, 2.5),
    design = "
    AB
    AC
    "
    )
fig_S1
  
```

```{r}

ggsave(
  plot = fig_S1, filename = str_c(fig_out_dir(), "FigS1.pdf"), 
  width = 6.5, height = 3, scale = 3, device = cairo_pdf
  )

```



## [deprecated] Supplementary Figure 2 (before revisisons)


```{r}

load(str_c(fig_out_dir(), "supplementary_ggplots/S2A.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S2B.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S2C.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S2D.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S2E.RData"), verbose = TRUE)


```

```{r}
#| fig-width: 13
#| fig-height: 16


fig_S2 <- 
  g_S2A + g_S2B +
  free(g_S2C + scale_x_continuous("Relative abundance", labels = scales::percent)) +
  g_S2D +
  g_S2E +
  
  plot_layout(
    heights = c(2,1,2,1.5),
    design = "
    AB
    CC
    DD
    EE
    "
    ) +
  plot_annotation(tag_levels = "A")
fig_S2
  
```

```{r}

ggsave(
  plot = fig_S2, filename = str_c(fig_out_dir(), "FigS2.pdf"), 
  width = 6.5, height = 8, scale = 2, device = cairo_pdf
  )

```

## Supplementary Figure 1 (final version, combining S1+S2)


```{r}
#| fig-width: 18
#| fig-height: 24

fig_S1 <- 
  free(g_S1A + labs(tag = "A")) + # + theme(legend.position = "bottom")
  g_S1B + labs(tag = "B") +
  g_S1C + labs(tag = "C") + 
  g_S2A + labs(tag = "D") + 
  g_S2B + labs(tag = "E") +
  free(g_S2C + scale_x_continuous("Relative abundance", labels = scales::percent)) + labs(tag = "F") +
  g_S2D + labs(tag = "G") + # G
  g_S2E + labs(tag = "H") + # H
  
  plot_layout(
    widths = c(1.6, 1, 1.4, 1.4),
    heights = c(1, 1, 2, 2, 1.5),
    design = "
    ABBB
    ACCC
    DDEF
    GGGG
    HHHH
    "
    )
fig_S1
  
```




```{r}

ggsave(
  plot = fig_S1, filename = str_c(fig_out_dir(), "FigS1.pdf"), 
  width = 6.5, height = 9, scale = 2.5, device = cairo_pdf
  )


```



## Supplementary Figure 3 (previously S4)


```{r}

load(str_c(fig_out_dir(), "supplementary_ggplots/S4A.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S4B.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S4C.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S4D.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S4EG.RData"), verbose = TRUE)

```

```{r}
#| fig-width: 13
#| fig-height: 15

set.seed(1)
fig_S3 <- 
  g_S4A + labs(caption = "") +  # A
  g_S4B + coord_fixed() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + # B
  g_S4C + theme(legend.position = "top") +
  g_S4D + theme(legend.position = "top") +
  g_S4E + g_S4F + g_S4G +
  plot_layout(
    heights = c(2,1,1.5),
    design = "
    AABBB
    CDDDD
    EFFGG
    "
    ) +
  plot_annotation(tag_levels = "A")
fig_S3
  
```

```{r}

set.seed(1)
ggsave(
  plot = fig_S3, filename = str_c(fig_out_dir(), "FigS3.pdf"), 
  width = 6.5, height = 8, scale = 2, device = cairo_pdf
  )

```



## Supplementary Figure 4 (previously S5)


```{r}

load(str_c(fig_out_dir(), "supplementary_ggplots/S5A.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S5B.RData"), verbose = TRUE)


```

```{r}
#| fig-width: 13
#| fig-height: 7

fig_S4 <- 
  g_S5A + g_S5B +
  plot_layout(ncol = 1) +
  plot_annotation(tag_levels = "A")
fig_S4
  
```

```{r}

ggsave(
  plot = fig_S4, filename = str_c(fig_out_dir(), "FigS4.pdf"), 
  width = 5, height = 3.5, scale = 2, device = cairo_pdf
  )

```




## [deprecated] Supplementary Figure 6 (before revisions)

```{r}

load(str_c(fig_out_dir(), "supplementary_ggplots/S6A.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S6B.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S6C.RData"), verbose = TRUE)

```


```{r}
#| fig-width: 12
#| fig-height: 12


fig_S6 <- 
  g_6A + labs(tag = "A") + 
  g_6B + 
  g_6C +
  plot_layout(
    guides = "collect", nrow = 3, heights = c(1.1,1,1)
  ) 

fig_S6

```

```{r}
#| eval: false

ggsave(
  plot = fig_S6, filename = str_c(fig_out_dir(), "FigS6.pdf"), 
  width = 6.5, height = 6.5, scale = 2, device = cairo_pdf
  )

```



## [deprecated] Supplementary Figure 7 (before revisions)

```{r}

load(str_c(fig_out_dir(), "supplementary_ggplots/S7A.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S7B.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S7C.RData"), verbose = TRUE)
load(str_c(fig_out_dir(), "supplementary_ggplots/S7D.RData"), verbose = TRUE)


```


```{r}
#| fig-width: 12
#| fig-height: 16


fig_S7 <- 
  free(g_S7A) +
  g_S7B + 
  g_S7C + ylab("") + theme(axis.text.y = element_blank()) +
  g_S7D + ylab("") + theme(axis.text.y = element_blank()) +
  plot_layout(
    heights = c(5,1),
    design = "
    AAA
    BCD
    "
  ) +
  plot_annotation(tag_levels = "A")

fig_S7

```

```{r}
#| eval: false

ggsave(
  plot = fig_S7, filename = str_c(fig_out_dir(), "FigS7.pdf"), 
  width = 6.5, height = 8, scale = 2.5, device = cairo_pdf
  )

```


## Supplementary Figure 5



```{r}
#| eval: false
#| fig-width: 18
#| fig-height: 26


fig_S5 <- 
  free(g_6A) + labs(tag = "A") + 
  free(g_6B) + 
  free(g_6C) +
  
  free(g_S7A) + # D
  g_S7B +  # E
  g_S7C + ylab("") + theme(axis.text.y = element_blank()) + # F
  g_S7D + ylab("") + theme(axis.text.y = element_blank()) + # G
  plot_layout(
    guides = "collect",
    heights = c(1.1, 1, 1, 5, 1),
    design = "
    AAA
    BBB
    CCC
    DDD
    EFG
    "
  ) +
  plot_annotation(tag_levels = "A")

fig_S5


```


```{r}
#| fig-width: 18
#| fig-height: 26


fig_S5 <- 
  free(g_6A) + labs(tag = "A") + 
  free(g_6B) + 
  free(g_6C) +
  
  free(g_S7A) + # D
  g_S7B +  # E
  g_S7C + ylab("") + theme(axis.text.y = element_blank()) + # F
  g_S7D + ylab("") + theme(axis.text.y = element_blank()) + # G
  plot_layout(
    guides = "collect",
    heights = c(1.5, 1.7, 5, 1),
    design = "
    AAAAAA
    BBBCCC
    DDDDDD
    EEFFGG
    "
  ) +
  plot_annotation(tag_levels = "A")

fig_S5


```

```{r}
#| eval: false

ggsave(
  plot = fig_S5, filename = str_c(fig_out_dir(), "FigS5.pdf"), 
  width = 6.5, height = 9.5, scale = 2.5, device = cairo_pdf
  )

```



## Supplementary Figure 6 (previously S8)


```{r}

load(str_c(fig_out_dir(), "supplementary_ggplots/S8.RData"), verbose = TRUE)


```


```{r}
#| fig-width: 13
#| fig-height: 12

fig_S6 <- g_S8
fig_S6

```

```{r}

ggsave(
  plot = fig_S6, filename = str_c(fig_out_dir(), "FigS6.pdf"), 
  width = 6.5, height = 6, scale = 2, device = cairo_pdf
  )

```


## Supplementary Figure 7 (previously S9)


```{r}

load(str_c(fig_out_dir(), "supplementary_ggplots/S9.RData"), verbose = TRUE)


```


```{r}
#| fig-width: 13
#| fig-height: 12

fig_S7 <- g_S9
fig_S7

```

```{r}

ggsave(
  plot = fig_S7, filename = str_c(fig_out_dir(), "FigS7.pdf"), 
  width = 6.5, height = 7, scale = 2, device = cairo_pdf
  )

```


# Data exports

```{r}
#| eval: false
source("../03_analyses/01_primary_and_secondary_endpoints/R/compute_BC_between_visits.R")

data_for_Laurel <- 
  mae@colData |> 
  as_tibble() |> 
  filter(!is.na(ARM)) |> 
  select(UID = Barcode, USUBJID, SITENAME, ARM, AVISITN, DAY, BV, LOAD, LCRISP, CTV05, LACTO) |> 
  left_join(
    compute_BC_between_successive_visits(mae, assayname = "ASV_16S_filtered_p", v = c(0:4,7))  |> 
      select(USUBJID, UID = Barcode_v1, BC_to_next_visit = BC, next_visit_AVISITN = v2),
    by = join_by(UID, USUBJID)
  ) |> 
  left_join(
    get_assay_wide_format(mae, "c_topics_16S_8", add_colData = FALSE) |> 
      dplyr::rename(UID = Barcode) |> 
      mutate(
        topic_prop = 
          assay |> 
          set_colnames(
            mae[["c_topics_16S_8"]] |> 
              rowData() |> 
              as.data.frame() |> 
              pull(topic_label))
        ) |> 
      dplyr::select(UID, topic_prop),
    by = join_by(UID)
  ) |> 
  filter(!is.na(topic_prop$`L. crispatus`)) |> 
  arrange(USUBJID, AVISITN) |> 
  unnest(topic_prop)

write_csv(
  data_for_Laurel, 
  file = 
    str_c(
      data_dir(), "06_exports/", 
      today() |> str_remove_all("-"), "_data_for_Laurel.csv"
      )
  )

```






```{r}

zenodo_data_dir <- str_c(data_dir(), "06_exports/Zenodo/")
if(!fs::dir_exists(zenodo_data_dir)) fs::dir_create(zenodo_data_dir)

write_rds(
  mae, 
  file = 
    str_c(
      zenodo_data_dir, "US_LactinV_multiomics_mae_", 
      today() |> str_remove_all("-"),
      ".rds"
    )
)



```

```{r}

source("export_mae_to_csv.R")
library(SummarizedExperiment)

export_mae_to_csv(
  mae, # mae[,,c(1,10, 24)], 
  dir = 
    str_c(
      zenodo_data_dir, "US_LactinV_multiomics_mae_", 
      today() |> str_remove_all("-"),
      "_as_csv_files/"
    )
)

```



# Session Info

<button class="accordion-button" type="button" data-bs-toggle="collapse" data-bs-target="#collapseOne" >Click this text for obtaining the R session information</button><div id="collapseOne" class="accordion-collapse collapse"><div>

```{r}
sessionInfo()
```

</div></div>
