---
title: "Are '-omics' datasets balanced at V0 and V1"
format: html
editor: visual
---

# Checking for arm differences at baseline?

```{r}
#| echo: false
#| warning: false
#| cache: false

library(tidyverse)
library(patchwork)
library(gtsummary)

theme_set(theme_light())


tmp <- fs::dir_map("R", source)
tmp <- fs::dir_map("../../R/", source)


```



```{r}

mae <- load_latest_mae(str_c(data_dir(), "03_augmented_mae/"))
clin <- MultiAssayExperiment::colData(mae) |> data.frame() |> as_tibble()

```

In this documents, we perform a series of check to ensure that the two groups (Lactin-V and placebo recipients) have similar characteristics at baseline.

For each of the domains (demographic variables, microbiota composition, cytokine/chemokine concentrations, behavioral variable), we check that balance for participants included for evaluating the **primary outcome** (= *L. crispatus* ≥ 50%) at **week 12**. We also repeat that analysis for participants with 16S data at **week 24**.

```{r}

week12_participants <- clin$USUBJID[!is.na(clin$AVISITN) & (clin$AVISITN == 4)]
week24_participants <- clin$USUBJID[!is.na(clin$AVISITN) & (clin$AVISITN == 7)]

```

## Demographics ("Table 1")

```{r}

dem_data <- 
  MultiAssayExperiment::colData(mae[,,"ASV_16S"]) |> 
  as_tibble() |> 
  select(USUBJID, ARM, AGE, RACEGR2, ETHNIC,  EDULVL, N_PAST_BV) |> 
  dplyr::rename(
    Age = AGE,
    `Self-declared race` = RACEGR2,
    Ethnicity = ETHNIC,
    Education = EDULVL,
    `Nb of past BV episodes` = N_PAST_BV
  )

```

```{r}

dem_data |> 
  filter(USUBJID %in% week12_participants) |> 
  distinct() |> 
  select(-USUBJID) |> 
  tbl_summary(by = ARM) |> 
  add_p() |> 
  modify_caption(caption = "Demographic variables of participants with microbiota composition data at week 12")


```

```{r}

dem_data |> 
  filter(USUBJID %in% week24_participants) |> 
  distinct() |> 
  select(-USUBJID) |> 
  tbl_summary(by = ARM) |> 
  add_p() |> 
  modify_caption(caption = "Demographic variables of participants with microbiota composition data at week 24")

```

## -omics at baseline (pre- and post-MTZ)

### 16S rRNA sequencing data

**Species composition**

```{r}

species <- MultiAssayExperiment::assay(mae, "tax_16S_p") %>% t()

```

We first compare the species composition at the first two visits (pre- and post-MTZ) for participants included for primary outcome at week 12.

```{r}
#| fig-width: 13
#| fig-height: 4

PCoA_baseline(
  mae = mae, assay_name = "tax_16S_p", 
  visit = 0, included_participants = week12_participants
  )

PCoA_baseline(
  mae = mae, assay_name = "tax_16S_p", 
  visit = 1, included_participants = week12_participants
  )

```

And for participants with microbiota composition at week 24.

```{r}
#| fig-width: 13
#| fig-height: 4

PCoA_baseline(
  mae = mae, assay_name = "tax_16S_p", 
  visit = 0, included_participants = week24_participants
  )

PCoA_baseline(
  mae = mae, assay_name = "tax_16S_p", 
  visit = 1, included_participants = week24_participants
  )

```

**ASV composition**

```{r}
#| fig-width: 13
#| fig-height: 4

set.seed(12345)

g1 <- 
PCoA_baseline(
  mae = mae, assay_name = "ASV_16S_filtered", 
  visit = 0, included_participants = week12_participants
)

g2 <- 
PCoA_baseline(
  mae = mae, assay_name = "ASV_16S_filtered", 
  visit = 1, included_participants = week12_participants
)

g1

g2


```


```{r}


  ggsave(g1, filename = str_c(fig_out_dir(), "S1B.pdf"), width = 13, height = 4, device = cairo_pdf)

```

```{r}

  ggsave(g2, filename = str_c(fig_out_dir(), "S1C.pdf"), width = 13, height = 4, device = cairo_pdf)

```



And for participants with microbiota composition at week 24.


```{r}
#| fig-width: 13
#| fig-height: 4

set.seed(12345)

PCoA_baseline(
  mae = mae, assay_name = "ASV_16S_filtered", 
  visit = 0, included_participants = week24_participants
)

PCoA_baseline(
  mae = mae, assay_name = "ASV_16S_filtered", 
  visit = 1, included_participants = week24_participants
)


```

**Topic composition**

```{r}

gammas <- 
  get_assay_long_format(
    mae, assayname = "c_topics_16S_8", add_rowData = TRUE,
    feature_name = "topic", values_name = "prop"
    )

```

```{r}
#| fig-height: 4
#| fig-width: 10

plot_topic_composition_1_visit(gammas, visit = 0) +
  labs(subtitle = "Screening visit")


```


```{r}

test_v0 <- test_topic_comp_by_ARM(gammas, visit = 0)
test_v0$full %>% summary()

anova(test_v0$full, test_v0$null)

```

No significant differences in microbiota composition at baseline when characterizing microbiota with topic composition.

```{r}
#| fig-height: 4
#| fig-width: 10

plot_topic_composition_1_visit(gammas, visit = 1) +
  labs(subtitle = "Randomization visit")


```

```{r}

test_v1 <- test_topic_comp_by_ARM(gammas, visit = 1)
test_v1$full |> summary()

anova(test_v1$full, test_v1$null)

```

```{r}
#| fig-width: 10
#| fig-height: 5

gammas |> 
  filter(AVISITN %in% 0:1, EOSSTT == "COMPLETED", !is.na(ARM)) |> 
  ggplot() +
  aes(x = topic_label, y = prop, fill = ARM) +
  geom_boxplot(outlier.size = 0.5) +
  facet_grid(. ~ ifelse(AVISITN == 0, "Screening (pre-MTZ)", "Randomization (post-MTZ)") |> fct_inorder()) +
  scale_fill_manual("", values = get_arm_colors()) +
  theme(
    strip.background = element_rect(fill = "gray", color = NA),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
    ) +
  xlab("Topic")

```

At the randomization visit, there are slightly more participants with subcommunity IV-B in the Placebo arm than in the Lactin-V arm.

**CST distribution**

```{r}

CSTs <- 
  get_assay_wide_format(mae, "CST") |>  
  unnest(cols = c(assay)) |> 
  mutate(Visit = str_c("Visit ", AVISITN))
  
```

For participants with microbiota data at week 12

```{r}

week12_CSTs <- 
  CSTs |> filter(AVISITN %in% 0:1, USUBJID %in% week12_participants)

 week12_CSTs |> 
  ggplot(aes(x = subCST, fill = CST)) +
  geom_bar() +
  facet_grid(ARM ~ Visit, scales = "free_y") +
  ylab("Nb of participants") +
   scale_fill_manual(values = get_topic_colors(week12_CSTs$CST |> unique() |> sort()))


week12_CSTs_V0 <- 
  week12_CSTs |> 
  filter(AVISITN == 0) |>  
  select(subCST, ARM)

(
  V0_chisq <- 
    chisq.test(table(week12_CSTs_V0), simulate.p.value = TRUE)
)


week12_CSTs_V1 <- 
  week12_CSTs |> 
  filter(AVISITN == 1) |>  
  select(subCST, ARM)

(
  V1_chisq <- 
    chisq.test(table(week12_CSTs_V1), simulate.p.value = TRUE)
)



```

For participants with microbiota data at week 24

```{r}

week24_CSTs <- 
  CSTs |> filter(AVISITN %in% 0:1, USUBJID %in% week24_participants)

 week24_CSTs |> 
  ggplot(aes(x = subCST, fill = CST)) +
  geom_bar() +
  facet_grid(ARM ~ Visit, scales = "free_y") +
  ylab("Nb of participants") +
   scale_fill_manual(values = get_topic_colors(week12_CSTs$CST |> unique() |> sort()))

week24_CSTs_V0 <- 
  week24_CSTs |> 
  filter(AVISITN == 0) |>  
  select(subCST, ARM)

(
  V0_chisq <- 
    chisq.test(table(week24_CSTs_V0), simulate.p.value = TRUE)
)

week24_CSTs_V1 <- 
  week24_CSTs |> 
  filter(AVISITN == 1) |>  
  select(subCST, ARM)

(
  V1_chisq <- 
    chisq.test(table(week24_CSTs_V1), simulate.p.value = TRUE)
)

```
No differences at baseline, but some differences post-MTZ, at randomization.




